---
title: "Lab 2"
author: "Kira Archipov"
date: "2023-01-19"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(palmerpenguins)
library(AICcmodavg)
library(equatiomatic)
```

### Predicting penguin mass 
We want to create a model we can use in the field to quickly and easily estimate a penguin's mass, based on the subset of data in the `palmerpenguins` package.

```{r}
penguins_clean <- penguins %>%
  drop_na() %>%
  rename(mass = body_mass_g,
         bill_l = bill_length_mm,
         bill_d = bill_depth_mm,
         flip_l = flipper_length_mm)

mdl1 <- lm(mass ~ bill_l + bill_d + flip_l + species + sex + island, data = penguins_clean)
```


### creating a formula to be used in model 1
```{r}
f1 <- mass ~ bill_l + bill_d + flip_l + species + sex + island

# we can take this a plug it in to out model in place of lm(), to clean up code 

mdl1 <- lm(f1, data = penguins_clean)
```


### New models + formulas 
```{r}
f2 <- mass ~ bill_l + bill_d + flip_l + species + sex 

mdl2 <- lm(f2, data = penguins_clean)

f3 <- mass ~ bill_d + flip_l + species + sex

mdl3 <- lm(f3, data = penguins_clean)

#using AIC to see which is better 
AIC(mdl1, mdl2, mdl3)
#model 2 has lower AIC, which is preferred 

#try it using BIC, which rewards parsimony more strongly 
BIC(mdl1, mdl2, mdl3)
```



```{r}
AICcmodavg::AICc(mdl1)
AICcmodavg::AICc(mdl2)
AICcmodavg::AICc(mdl3)
#correction works better for really small samples 

aictab(list(mdl1, mdl2, mdl3))
bictab(list(mdl1, mdl2, mdl3))

# if difference is <2 its weak evidence that the smallest BIC model is better, 2-6 is moderately better, >6 strong evidence that the smallest BIC is the superior model 
```

### Compare models using K-fold cross validation - using manual entry 
1) break data into x folds (10 in this example)
2) take portion out 
```{r}
folds <- 10 

fold_vec <- rep(1:folds, length.out = nrow(penguins_clean))

#starting with exact same points of random # generator, would help with replication, only need to set it once and then it will be for all other random calls 
set.seed(42)

penguins_fold <- penguins_clean %>% 
  mutate(group = sample(fold_vec, size = n(), replace = FALSE))

table(penguins_fold$group)

test_df <- penguins_fold %>% 
  filter(group == 1)
train_df <- penguins_fold %>% 
  filter(group != 1)

```


### Creating a function comparing 2 vectors 

```{r}
calc_rmse <- function(x, y) {
  rmse <- (x - y)^2 %>% 
    mean() %>% 
    sqrt()
  return(rmse)
}
```



```{r}
training_mdl1 <- lm(f1, data = train_df)

training_mdl2 <- lm(f2, data = train_df)

training_mdl3 <- lm(f3, data = train_df)

#see how well these predict in the test data frame (the group that was held aside)

predict_test <- test_df %>% 
  mutate(model1 = predict(training_mdl1, test_df), 
         model2 = predict(training_mdl2, test_df), 
         model3 = predict(training_mdl3, test_df))

# how far away are these predicted values from the known values, how well are we predicting 
rmse_predict_test <- predict_test %>% 
  summarize(rmse_mdl1 = calc_rmse(model1, mass), 
            rmse_mdl2 = calc_rmse(model2, mass), 
            rmse_mdl3 = calc_rmse(model3, mass))

#this is one string of the marble example !!!
```


### lets iterate!

```{r}
#create an empty data frame
rmse_df <- data.frame()

#foreloops!

for(i in 1:folds) {
  kfold_test_df <- penguins_fold %>% 
    filter(group == i)
  kfold_train_df <- penguins_fold %>% 
    filter(group != i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  kfold_mdl3 <- lm(f3, data = kfold_train_df)
  
  kfold_predict_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df), 
           mdl2 = predict(kfold_mdl2, .),
           mdl3 = predict(kfold_mdl3, .))
  kfold_rmse_df <- kfold_predict_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, mass), 
              rmse_mdl2 = calc_rmse(mdl2, mass),
              rmse_mdl3 = calc_rmse(mdl3, mass), 
              test_gp = i)
  rmse_df <- bind_rows (rmse_df, kfold_rmse_df)
}

```








